---
title: "HW6 PR8"
author: "Yusuf Sina Öztürk - Ahmet Buğra Taksuk - Ahmet Tabakoğlu"
date: "7/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, libraries, include=F}
library(tidyverse)
library(lubridate)
library(zoo)
library(ggplot2)
library(data.table)
library(dplyr)
library(forecast)
library(miscTools)
library(stats)
library(GGally)
```

### PRODUCT ID: 4066298

For the product 2, I followed the same steps of product 1. Therefore, I am going to skip the descriptive comments and keep the related ones only.

### TASK 1 - Analyzing Seasonality


### 1 ) Importing and Manipulating the data



```{r,1}
pr8 = read.csv("alldata_item8.csv")
pr8 <- as.data.table(pr8)
pr8 <- pr8[,-c("X","w_day")]
```

```{r,2,warning=F}
pr8 <- mutate(pr8, event_date = ymd(event_date)) # converting event date into datetime object
pr8[, Month:=as.numeric(lubridate::month(event_date,label=F))] #adding month information as a numeric variable 
pr8[, Day:=as.numeric(lubridate::wday(event_date,label=F))] #adding day information as a numeric variable 
head(pr8)
```


```{r,3}
sold <- data.table(event_date =pr8$event_date,
                   sold_count = pr8$sold_count)
head(sold)
```

### Visualizations

```{r,4, fig.width=10}
boxplot(sold$sold_count)
```

```{r,5,fig.width=10}
ggplot(sold, aes(x = event_date)) + 
  geom_line(aes(y = sold_count)) +  ggtitle("Product 85004 Sold Amount") +
  xlab("Date") + ylab("Amount Sold")

```

#### ACF Plot

```{r,6,fig.width=10}
acf(sold$sold_count)
```

#### PACF Plot

```{r,7,fig.width=10}
pacf(sold$sold_count)
```

### 3 ) Decomposing the data

#### Weekly Seasonality

```{r,8,fig.width=10}
soldts <- ts(rev(pr8$sold_count),  freq = 7, start= c(1,1))
resultweekdec <- decompose(soldts,type= "additive")
plot(resultweekdec)
```
#### Monthly  Seasonality

```{r,9,fig.width=10}

soldtsmonth <- ts(rev(pr8$sold_count),  freq = 30, start= c(1,1))
resultmonthdec <- decompose(soldtsmonth,type= "additive")
plot(resultmonthdec)
```

#### Random term after decomposing weekly

```{r,10,fig.width=10}
plot(resultweekdec$random)
title("Random Term of Weekly Decomposed Data")
```
#### Random term after decomposing monthly

```{r,11,fig.width=10}
plot(resultmonthdec$random)
title("Random Term of Monthly  Decomposed Data")
```

#### Conclusion

Both the monthly and the weekly decompositon seems to have a significant seasonal component but the trend variables are not significant. We might use weekly decompositon as the random term of weekly decomposition seems more like white noise series.

```{r,12,fig.width=10}
random = resultweekdec$random
trend = resultweekdec$trend
season = resultweekdec$seasonal
```

### Task 2- Fitting data into ARIMA Model

#### 1 ) Deciding on the parameters

To decide on the parameters of the ARIMA model, I checked the ACF and PACF plots of the data. ACF plot has exponentially decaying and sinusodial pattern and PACF function has a sudden drop at lag 1 and no significant value after lag 1. So the model suggest a AR(1) model. But I am going to check the neighborhood to find the best model.


```{r,13,fig.width=10}
acf(random, na.action = na.pass)
```

```{r,12213,fig.width=10}
pacf(random, na.action = na.pass)
```
```{r,14,fig.width=10}
model <- arima(random, order= c(1,0,0))
AIC(model)

model <- arima(random, order= c(2,0,0))
AIC(model)

model <- arima(random, order= c(3,0,0))
AIC(model)

model <- arima(random, order= c(4,0,0))
AIC(model)

model <- arima(random, order= c(1,0,1))
AIC(model)

model <- arima(random, order= c(2,0,1))
AIC(model)

model <- arima(random, order= c(3,0,1))
AIC(model)

model <- arima(random, order= c(4,0,1))
AIC(model)

model <- arima(random, order= c(3,0,2))
AIC(model)

model <- arima(random, order= c(2,0,2))
AIC(model)

model <- arima(random, order= c(1,0,2))
AIC(model)
```


So the best model we came up with is the one with (3,0,2) three autoregressive term and two moving average term.

#### 2 ) Fitting the Model

```{r,15,fig.width=10}
model <- arima(random, order= c(3,0,2))
summary(model)
```

#### Residuals after fitting

```{r,16,fig.width=10}
plot(model$residuals)
title("Residuals")
```

```{r,17,fig.width=10}
modelfit <- random - model$residuals
fitted <- modelfit+trend+season

plot(soldts)
points(fitted, type= "l", col = 2)

```
The model seems to have a good fit since zero mean assumption and constant varinace assumption seems to be held but it can be improved adding extra regressors to the model.

### Task 3 and 4 - Adding external regressors to the model

#### 1 ) Searcing for regressors

In this part, I am going to look for variables that might be correlated to sold amount and can be used as regressors in the ARIMAX model. To do this, first I am going to use pairplots. But some of the data in the columns are lost for so many data points. So only columns I can use for this, "Basket Count" , "Category Sold" , Category Visits" and "Category Favored". I checked the relation of each variable with the sold count. From the visualizations, Basket Count, Category Favored and Category Sold seems to have a correlation, so I am going to add those variables to the model.

```{r,18,fig.width=10, warning=F}
ggpairs(pr8, columns = c(4,7,8,10,13 ))
```

#### 2 ) Adding regressors to the ARIMA model

```{r,19,fig.width=10, warning=F}
traindata <- sold[-c(1:7),]
head(traindata)
testdata <- sold[c(1:7),]
head(testdata)
```

```{r,20,fig.width=10, warning=F}
regressors <- pr8[-c(1:7), c(7,8,13)]
head(regressors)
```

```{r,21,fig.width=10, warning=F}
traindatats <- ts(rev(traindata$sold_count),frequency = 7, start = c(1,1))
resultdec <- decompose(traindatats,type= "additive")
trend = resultdec$trend
season = resultdec$seasonal
random = resultdec$random 
plot(resultdec)
```
#### Without regressors

```{r,22,fig.width=10, warning=F}
model <- arima(random, order= c(3,0,2))
summary(model)
```
#### With external regressor

```{r,23,fig.width=10, warning=F}
model <- arima(random, order= c(3,0,2), xreg = regressors)
summary(model)
```

First AR term has lost its significancy . However I am going to forecast with this model.

#### Forecasting and model performance measures

```{r,25,fig.width=10, warning=F}
newregmatrix <- pr8[c(1:10), c(7,8,13)]
  
model.forecast <- predict(model, n.ahead = 10, newxreg = newregmatrix)$pred

last.trend.value <-tail(resultdec$trend[!is.na(resultdec$trend)],1)
seasonality <- resultdec$seasonal[370:379]

forecast_normalized <- model.forecast+last.trend.value+seasonality
forecast_normalized= ts(forecast_normalized, frequency = 7, start=c(55,3))
```

```{r,26,fig.width=10, warning=F}
testdata <- ts(rev(testdata$sold_count), frequency = 7, start=c(55,3))

plot(testdata)
points(forecast_normalized,type= "l", col = 2)

```

```{r,27,fig.width=10, warning=F}
modelfit <- random - model$residuals

plot(random)
points(modelfit, type= "l", col = 2)

```
```{r,28,fig.width=10, warning=F}
plot(model$residuals)

```

#### Conclusion

##### Overall looking the model, zero mean and independency of the residuals assumptions seem to be held by looking at the residuals. For this product, I might use ARIMAX model rather than LM model. But the at some points, peaks should be excluded.
